Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-21 cav2017/cav2017.tex
More reductions

\todo{Dario, can you please shorten this?}

@@

Let $\mathbb{R} \langle I,F \rangle$ be a fixed point domain with $I$ bits representing the integer part and $F$ bits
representing the decimal part. 
The following approximation errors will arise:
\begin{enumerate}
\item {\bf Truncation:}
Let $\mathcal{F}\langle I,F \rangle(\cdot) : \mathbb{R} \rightarrow \mathbb{R}\langle I,F \rangle, \tilde x=\mathcal{F}\langle I,F \rangle(x) = x-\delta : \delta=x\ \%_F\  c\tilde x$, where $\%_F$ is the modulus operation performed on the last bit of the representation ($2^{-F}$).
The number $\delta$ corresponds to the truncation error of the representation and it will propagate across operations.
\item {\bf Rounding:}
%% Every time an operation is performed on a number in the $\mathbb{R} \langle I,F,E \rangle$ domain, an error may be
%% introduced due to truncation of the result.
Let us define $c\langle i,f \rangle \in \mathbb{R}\langle I,F\rangle$, where $i \in [-2^{I-1}\ 2^{I-1}],f \in [0 1)$ are its assigned values. The smallest number that can be represented for any given exponent is $c_{min}=2^{-F}$.
The following errors appear in simple operations:
\begin{enumerate}
\item Addition/Substraction: in the case of fixed point arithmetic of numbers with the same domain, there are no rounding errors, otherwise the maximum error is given by the domain of the result $c\langle \cdot, F_1\rangle \pm c\langle \cdot , F_2\rangle= c\langle \cdot , F_3\rangle + \delta_3 :|\delta_3| \leq 2^{-F_3} \wedge F_3 \leq F_1,F_2$ 
\item Multiplication: $c\langle \cdot, F_1\rangle * c\langle \cdot,F_2\rangle=c\langle \cdot,F_3\rangle + \delta_3 : |\delta_3| \leq 2^{-F_3} \wedge F_3 \leq F_1,F_2$ 
\end{enumerate}
The bounds obtained for division are left to the reader as they will not be applied in this paper.

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Fixed word length effects on matrix multiplications}
\label{sec:cof_fwl}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
Thus far we have presented a verification procedure for a digitalized plant in a continuous space (\emph{ie} the
range of the quatizers is still $\mathbb{R}$. The next step we must reaize is that the controller and the observer
exist in a digital program, hence their domain is $\mathbb{R}\langle I,F\rangle$. This has two separate effects
on the model. If we recall equation \eqref{eq:to_of}, the matrices of the observer and the controller  are not necessarily those of the plant.
In fact we have $\mat{A},\mat{B},\vec{x} \in \mathbb{R}^*$, $\mat{K}_d,\vec{u} \in \mathbb{R}\langle I,F\rangle^*$
Where $*$ represents the appropriate dimension for each matrix/vector.
Unfortunately this means that although combined matrices such as $\mat{A}_t$ are in the reals, they suffer from
the fixed word restrictions in their composing matrices.
The second effect corresponds to the rounding errors in the calculation of $\hat{x}_{k+1}, u_k$. These will introduce a new source of quantiztion noise which we may denote $\vec{\nu}_3 \in [-q_3\ \ q_3]$.
The estimation of $\hat{\vec{x}}$ corresponds to the sum of three matrix-vector multiplications. Each of these has a coefficient-wise error bound $c_{min}$. Hence the overall $\overline{\delta}_{\hat{x}}=3nc_{min}$. The error in $\vec{u}$ corresponds of the multiplication of $\mat{K}_d\hat{\vec{x}}$, hence $q_3=\overline{\delta}_u=nc_{min}+\sum_{k=1}^nc_k\overline{\delta}_{\hat{x}}=(3|\mat{K}|_1+1)n2^{-F}$

>>>>

@@

Let $\mathbb{R} \langle I,F \rangle$ be a fixed point domain with $I$ bits representing the integer part and $F$ bits
representing the decimal part. The smallest number that can be represented in this domain is $c_m=2^{-F}$.
The following approximation errors will arise:
\begin{enumerate}
\item {\bf Truncation:}
Let $\mathcal{F}\langle I,F \rangle(\cdot) : \mathbb{R} \rightarrow \mathbb{R}\langle I,F \rangle, \tilde x=\mathcal{F}\langle I,F \rangle(x) = x-\delta_T : \delta_T=x\ \%_{c_m}\ \tilde x$, where $\%_{c_m}$ is the modulus operation performed wrt the last bit.
$\delta_T$ is the truncation error and it will propagate across operations.
\item {\bf Rounding:}
The following errors appear in simple operations:
\begin{enumerate}
\item Addition/Substraction: these operations only propagate errors coming from truncation of the operands.
$c_1\pm c_2= c_3 + \delta_3 :|\delta_3| \leq |\delta_{T1}| + |\delta_{T2}|$ 
\item Multiplication: $c_1 * c_2= c_3 + \delta_3 :|\delta_3| \leq |\delta_{T1}*c_2| + |\delta_{T2}*c_1| + c_m$ 
\item Division will not be applied in these implementations.
\end{enumerate}

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Fixed word length effects on matrix multiplications}
\label{sec:cof_fwl}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
Thus far we have presented a verification procedure for a digitalized plant in a continuous space (\emph{ie} the
range of the quatizers is still $\mathbb{R}$. The next step we must reaize is that the controller 
exists in a digital program, hence its domain is $\mathbb{R}\langle I,F\rangle$. We have $\mat{A},\mat{B},\vec{x} \in \mathbb{R}^*$, $\mat{K}_d,\vec{u} \in \mathbb{R}\langle I,F\rangle^*$
Where $*$ represents the appropriate dimension for each matrix/vector.
Unfortunately this means that although combined matrices such as $\mat{A}_t$ are in the reals, they suffer from
the fixed word restrictions in their composing matrices.
The second effect corresponds to the rounding errors in the calculation of $\hat{x}_{k+1}, u_k$. These will introduce a new source of quantiztion noise which we may denote $\vec{\nu}_3 \in [-q_3\ \ q_3]$.
The estimation of $\hat{\vec{x}}$ corresponds to the sum of three matrix-vector multiplications. Each of these has a coefficient-wise error bound $c_{min}$. Hence the overall $\overline{\delta}_{\hat{x}}=3nc_{min}$. The error in $\vec{u}$ corresponds of the multiplication of $\mat{K}_d\hat{\vec{x}}$, hence $q_3=\overline{\delta}_u=nc_{min}+\sum_{k=1}^nc_k\overline{\delta}_{\hat{x}}=(3|\mat{K}|_1+1)n2^{-F}$
